---
execute:
    freeze: false
---

# PARE model

## Load data, functions, and packages

```{r}
#| label: setup
#| message: false
#| code-fold: true

library(dplyr)
library(sf)
library(spdep)
library(extRemes)
library(ggplot2)
library(patchwork)

thresh = 253

stations <- read.csv("Data/station_info.csv")
stat_nos <- stations[,1]

source("Code/HelperFunctions_PARE/format_data.R")
source("Code/HelperFunctions_PARE/jitter_n_sym.R")
source("Code/HelperFunctions_PARE/get_sym_car_mat.R")
source("Code/HelperFunctions_PARE/get_par_car.R")
source("Code/HelperFunctions_PARE/get_se.R")
source("Code/HelperFunctions_PARE/pars_to_rl.R")
source("Code/HelperFunctions_PARE/rl_with_ci.R")
source("Code/HelperFunctions_PARE/make_varcov_mats_car.R")
source("Code/HelperFunctions_PARE/rl_ci_to_plot_vec.R")
source("Code/HelperFunctions_PARE/create_mat.R")
source("Code/HelperFunctions_PARE/win_rl_to_plot_dat.R")
source("Code/HelperFunctions_PARE/get_latex_table.R")
```

## Format time series of GPD fits

```{r}
#| label: load_gpd
#| eval: false

## load watershed data
ws_regs <- sf::read_sf("Data/watershed_region_updated")
st_crs(ws_regs) <-sf::st_crs(2278)

## window object generated in Fit_GPD.qmd
## Load rolling GPD fits (takes a bit)
window <- readRDS("Data/window_1day_dclust_updated.rds")  # list of lists (82 x 601) of GPD fits
window_CVM <- readRDS("Data/window_CVM_updated.rds")
window_AD <- readRDS("Data/window_AD_updated.rds")

## Extract windows (note varying # stations)
dat_win_22 <- format_data(window[[22]], stations, ws_regs, window_CVM[[22]]) # 34
dat_win_52 <- format_data(window[[52]], stations, ws_regs, window_CVM[[52]]) # 28
dat_win_82 <- format_data(window[[82]], stations, ws_regs, window_CVM[[82]]) # 149

## save values (faster than reading entire window[[]] object)
save(dat_win_22, file = "Data/dat_win_22")
save(dat_win_52, file = "Data/dat_win_52")
save(dat_win_82, file = "Data/dat_win_82")
```

```{r}
#| label: load_gpd_precomputed
#| echo: false
#| eval: true
#| message: false

load("Data/dat_win_22")
load("Data/dat_win_52")
load("Data/dat_win_82")
```

## Construct spatial weight matrices

::: {.callout-note collapse="true"}
## Generating the Hausdorff matrix

The `hausMat` function and additional documentation can be found in the [`hausdorff` Github repository](https://github.com/juliaSchedler/hausdorff).

```{r}
#| eval: false
#| label: hausMat

# (make sure ws_regs is projected first)
distMat <- hausMat(ws_regs, f1 = 0.5)
distMat/5280

# save output
# saveRDS(distMat, file = "~/Documents/GitHub/Spatial_Extreme_Value_Modeling/Data/hMat_med.rds")

```
:::

```{r}
#| label: weight-matrix
# read in previously computed Hausdorff matrix and convert units to miles
hMat <- readRDS(file = "Data/hMat_med.rds")
hMat_miles <- hMat/5280

## Create weight matrix for each window 
## Accounts for different numbers of stations (varying point-to-area structure)
## Jiters the values to give valid weight matrix
set.seed(24)
D_22 <- get_sym_car_mat(dat_win_22, hMat_miles)
D_52 <- get_sym_car_mat(dat_win_52, hMat_miles)
D_82 <- get_sym_car_mat(dat_win_82, hMat_miles)
```

## Fit models to each window for each parameter

To obtain region-level estimates of the GPD parameters, fit a conditional auto-regressive model for each GPD parameter for each window, a total of 9 models.

```{r}
#| label: fit_car_windows
#| eval: true

## window 22
## Fit models
car_shape_22 <- spatialreg::spautolm(shape ~ -1 + Reg1 + Reg2 + Reg3, 
                                     data = dat_win_22, family="CAR", 
                                     listw=mat2listw(1/D_22, style = "C"))
car_ln.scale_22 <- spatialreg::spautolm(log(scale) ~ -1 + Reg1 + Reg2 + Reg3, 
                                        data = dat_win_22, family="CAR", 
                                        listw=mat2listw(1/D_22,style = "C"))
car_rate_22 <- spatialreg::spautolm(rate ~ -1 + Reg1 + Reg2 + Reg3, 
                                    data = dat_win_22, family="CAR", 
                                    listw=mat2listw(1/D_22, style = "C"))
## extract parameter estimates
par_22 <- get_par_car(car_ln.scale_22, car_shape_22, car_rate_22)
se_22 <- get_se(car_ln.scale_22, car_shape_22, car_rate_22)[3,]
car_varcov_22 <- make_varcov_mats_car(car_ln.scale_22, car_shape_22)  # se of scale 8.5, 4, 9.25
st_devs_22 <- data.frame(Reg1 = c(sqrt(diag(car_varcov_22[[1]])), se_22[1]),
                         Reg2 = c(sqrt(diag(car_varcov_22[[2]])), se_22[2]), 
                         Reg3 = c(sqrt(diag(car_varcov_22[[3]])), se_22[3]))
row.names(st_devs_22) <- row.names(par_22)



## window 52
## Fit models
car_shape_52 <- spatialreg::spautolm(shape ~ -1 + Reg1 + Reg2 + Reg3, 
                                     data = dat_win_52, family="CAR", 
                                     listw=mat2listw(1/D_52, style = "C"))
car_ln.scale_52 <- spatialreg::spautolm(log(scale) ~ -1 + Reg1 + Reg2 + Reg3, 
                                        data = dat_win_52, family="CAR", 
                                        listw=mat2listw(1/D_52, style = "C"))
car_rate_52 <- spatialreg::spautolm(rate ~ -1 + Reg1 + Reg2 + Reg3, 
                                    data = dat_win_52, family="CAR", 
                                    listw=mat2listw(1/D_52, style = "C"))
## Extract parameter estimates
par_52 <- get_par_car(car_ln.scale_52, car_shape_52, car_rate_52)
se_52 <- get_se(car_ln.scale_52, car_shape_52, car_rate_52)[3,]
car_varcov_52 <- make_varcov_mats_car(car_ln.scale_52, car_shape_52)  # se of scale 8.5, 4, 9.25
st_devs_52 <- data.frame(Reg1 = c(sqrt(diag(car_varcov_52[[1]])), se_52[1]),
                         Reg2 = c(sqrt(diag(car_varcov_52[[2]])), se_52[2]), 
                         Reg3 = c(sqrt(diag(car_varcov_52[[3]])), se_52[3]))

## window 82
## fit models
car_shape_82 <- spatialreg::spautolm(shape ~ -1 + Reg1 + Reg2 + Reg3, 
                                     data = dat_win_82, family="CAR", 
                                     listw=mat2listw(1/D_82, style = "C"))
car_ln.scale_82 <- spatialreg::spautolm(log(scale) ~ -1 + Reg1 + Reg2 + Reg3, 
                                        data = dat_win_82, family="CAR", 
                                        listw=mat2listw(1/D_82, style = "C"))
car_rate_82 <- spatialreg::spautolm(rate ~ -1 + Reg1 + Reg2 + Reg3, 
                                    data = dat_win_82, family="CAR", 
                                    listw=mat2listw(1/D_82, style = "C"))
## Extract parameter estiamtes
par_82 <- get_par_car(car_ln.scale_82, car_shape_82, car_rate_82)
se_82 <- get_se(car_ln.scale_82, car_shape_82, car_rate_82)[3,]
car_varcov_82 <- make_varcov_mats_car(car_ln.scale_82, car_shape_82)  # se of scale 8.5, 4, 9.25
st_devs_82 <- data.frame(Reg1 = c(sqrt(diag(car_varcov_82[[1]])), se_82[1]),
                         Reg2 = c(sqrt(diag(car_varcov_82[[2]])), se_82[2]), 
                         Reg3 = c(sqrt(diag(car_varcov_82[[3]])), se_82[3]))

```

### Table of parameter estimates

```{r}
#| label: par-table
#| echo: false

## PARE part of Table 4
#get_latex_table(par_22, st_devs_22)
cat("PARE portion of Table S3 \n First 40 Years of Data (1921-1960) \n Parameters")
par_22
cat("Standard Errors")
st_devs_22
## PARE part of Table 5
#get_latex_table(par_52, st_devs_52)
cat("PARE portion of Table S4 \n Second 40 Years of Data (1951-1990) \n Parameters")
par_52
cat("Standard Errors")
st_devs_52
## PARE part of Table 2
#get_latex_table(par_82, st_devs_82)
cat("PARE portion of Table 2 \n Last 40 Years of Data (1981-2020) \n Parameters")
par_82
cat("Standard Errors")
st_devs_82

```

## Calculate return levels + CIs

```{r}
#| label: return-levels
### Estimates and CIs for visuals
car_rl_25_22  <- rl_with_ci(par_mat = par_22, 
                            varcov_list = car_varcov_22, 
                            return_period = 25, 
                            type = "ci", alpha = 0.05)/254
car_rl_100_22 <- rl_with_ci(par_22, car_varcov_22, 100, "ci")/254
car_rl_500_22 <- rl_with_ci(par_22, car_varcov_22, 500, "ci")/254

car_rl_25_52  <- rl_with_ci(par_52, car_varcov_52, 25, "ci")/254
car_rl_100_52 <- rl_with_ci(par_52, car_varcov_52, 100, "ci")/254
car_rl_500_52 <- rl_with_ci(par_52, car_varcov_52, 500, "ci")/254

car_rl_25_82  <- rl_with_ci(par_82, car_varcov_82, 25, "ci")/254
car_rl_100_82 <- rl_with_ci(par_82, car_varcov_82, 100, "ci")/254
car_rl_500_82 <- rl_with_ci(par_82, car_varcov_82, 500, "ci")/254

win_1 <- rl_ci_to_plot_vec(car_rl_25_22, car_rl_100_22, car_rl_500_22)
win_2 <- rl_ci_to_plot_vec(car_rl_25_52, car_rl_100_52, car_rl_500_52)
win_3 <- rl_ci_to_plot_vec(car_rl_25_82, car_rl_100_82, car_rl_500_82)

plot_dat <- win_rl_to_plot_dat(win_1, win_2, win_3)
reg1 <- plot_dat[[1]] # Region 1 data frame for plotting
reg2 <- plot_dat[[2]] 
reg3 <- plot_dat[[3]] 


```

### Visualizations

```{r}
#| label: PARE-visuals
## define horizontal limits for error bars
reg1$LB_X <- reg2$LB_X <- reg3$LB_X <- c(1957, 1987, 2017)
reg1$UB_X <- reg2$UB_X <- reg3$UB_X <- c(1963, 1993, 2023)

## legend 
g <- guides(fill = guide_legend(override.aes = list(color = c("black", "black", "black"),
                                                    shape = c(15, 16, 18), 
                                                    fill = c("black", "black", "black"),
                                                    linetype = c(1,1,1),
                                                    size = c(2,2,3)
                                                    )))

## Region 1
reg1_plot <- ggplot(data=reg1, aes(x=c(1960, 1990, 2020))) + 
    coord_cartesian(ylim = c(3, 41.5)) +
  ## 500 year
    geom_point(aes(y=RL_500, color="500-Year", shape = "500-Year"), size = 2) + 
    geom_errorbar(aes(ymin=LB_500, ymax=UB_500, col="500-Year", linetype = "500-Year"), width = 6)+ 
    geom_rect(aes(ymin=LB_500, ymax=UB_500,xmin = LB_X, xmax = UB_X, fill="500-Year"), alpha=0.1)+
  ## 100 year
    geom_point(aes(y=RL_100, color="100-Year", shape = "100-Year"), size = 2)+ 
    #geom_errorbar(aes(ymin=LB_100, ymax=UB_100, col="100-Year", linetype = "100-Year"), width = 6) +     
    geom_rect(aes(ymin=LB_100, ymax=UB_100,xmin = LB_X, xmax = UB_X, fill="100-Year"), alpha=0.1)+
  ##25 year
    geom_point(aes(y=RL_25, color="25-Year", shape = "25-Year"), size = 3) + 
    geom_errorbar(aes(ymin=LB_25, ymax=UB_25, col="25-Year", linetype = "25-Year"), width = 6) +
    geom_rect(aes(ymin=LB_25, ymax=UB_25,xmin = LB_X, xmax = UB_X, fill="25-Year"), alpha=0.1)+
    scale_x_continuous(breaks = c(1960, 1990,2020)) +
    scale_colour_manual(name = "Return Period", 
                        values = c('500-Year' = "darkblue", '100-Year' = "blue", '25-Year' = "skyblue"),
                        breaks = c('500-Year', '100-Year', '25-Year')) + 
    scale_linetype_manual(values = c("500-Year" = "solid", "100-Year" = "solid", "25-Year" = "solid"),
                          breaks = c('500-Year', '100-Year', '25-Year'), name = "Return Period") +
  scale_fill_manual(name = "Return Period", values =  c('500-Year' = "darkblue", '100-Year' = "blue", '25-Year' = "lightblue"), breaks = c('500-Year', '100-Year', '25-Year'))+
    scale_shape_manual(name = "Return Period", values = c('500-Year' = 15, '100-Year' = 16, '25-Year' = 18), breaks = c('500-Year', '100-Year', '25-Year')) +
     labs(x="", y="Return Level (in)", title="Region 1") + theme_minimal() +   theme(legend.position="none",  
        axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 15), 
        title = element_text(size = 15))

## Region 2
reg2_plot <- ggplot(data=reg2, aes(x=c(1960, 1990, 2020))) + 
    coord_cartesian(ylim = c(3, 41.5)) +
  ## 500 year
    geom_point(aes(y=RL_500, color="500-Year", shape = "500-Year"), size = 2) + 
    geom_errorbar(aes(ymin=LB_500, ymax=UB_500, col="500-Year", linetype = "500-Year"), width = 6)+ 
    geom_rect(aes(ymin=LB_500, ymax=UB_500,xmin = LB_X, xmax = UB_X, fill="500-Year"), alpha=0.1)+
  ## 100 year
    geom_point(aes(y=RL_100, color="100-Year", shape = "100-Year"), size = 2)+ 
    geom_errorbar(aes(ymin=LB_100, ymax=UB_100, col="100-Year", linetype = "100-Year"), width = 6) +     
    geom_rect(aes(ymin=LB_100, ymax=UB_100,xmin = LB_X, xmax = UB_X, fill="100-Year"), alpha=0.1)+
  ##25 year
    geom_point(aes(y=RL_25, color="25-Year", shape = "25-Year"), size = 3) + 
    geom_errorbar(aes(ymin=LB_25, ymax=UB_25, col="25-Year", linetype = "25-Year"), width = 6) +
    geom_rect(aes(ymin=LB_25, ymax=UB_25,xmin = LB_X, xmax = UB_X, fill="25-Year"), alpha=0.1)+
    scale_x_continuous(breaks = c(1960, 1990,2020)) +
    scale_colour_manual(name = "Return Period", 
                        values = c('500-Year' = "darkred", '100-Year' = "red", '25-Year' = "pink"),
                        breaks = c('500-Year', '100-Year', '25-Year')) + 
    scale_linetype_manual(values = c("500-Year" = "solid", "100-Year" = "solid", "25-Year" = "solid"),
                          breaks = c('500-Year', '100-Year', '25-Year'), name = "Return Period") +
  scale_fill_manual(name = "Return Period", values =  c('500-Year' = "darkred", '100-Year' = "red", '25-Year' = "pink"), breaks = c('500-Year', '100-Year', '25-Year'))+
    scale_shape_manual(name = "Return Period", values = c('500-Year' = 15, '100-Year' = 16, '25-Year' = 18), breaks = c('500-Year', '100-Year', '25-Year')) +
     labs(x="Last Year of 40-Year Window", y="", title="Region 2") + theme_minimal() + g +   theme(legend.position="bottom",  
        axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 15), 
        title = element_text(size = 15))



## Region 3
reg3_plot <- ggplot(data=reg3, aes(x=c(1960, 1990, 2020))) + 
    coord_cartesian(ylim = c(3, 41.5)) +
  ## 500 year
    geom_point(aes(y=RL_500, color="500-Year", shape = "500-Year"), size = 2) + 
    geom_errorbar(aes(ymin=LB_500, ymax=UB_500, col="500-Year", linetype = "500-Year"), width = 6)+ 
    geom_rect(aes(ymin=LB_500, ymax=UB_500,xmin = LB_X, xmax = UB_X, fill="500-Year"), alpha=0.1)+
  ## 100 year
    geom_point(aes(y=RL_100, color="100-Year", shape = "100-Year"), size = 2)+ 
    geom_errorbar(aes(ymin=LB_100, ymax=UB_100, col="100-Year", linetype = "100-Year"), width = 6) +     
    geom_rect(aes(ymin=LB_100, ymax=UB_100,xmin = LB_X, xmax = UB_X, fill="100-Year"), alpha=0.1)+
  ##25 year
    geom_point(aes(y=RL_25, color="25-Year", shape = "25-Year"), size = 3) + 
    geom_errorbar(aes(ymin=LB_25, ymax=UB_25, col="25-Year", linetype = "25-Year"), width = 6) +
    geom_rect(aes(ymin=LB_25, ymax=UB_25,xmin = LB_X, xmax = UB_X, fill="25-Year"), alpha=0.1)+
    scale_x_continuous(breaks = c(1960, 1990,2020)) +
    scale_colour_manual(name = "Return Period", 
                        values = c('500-Year' = "#256e3a", '100-Year' = "#5d916c", '25-Year' = "#44c96a"),
                        breaks = c('500-Year', '100-Year', '25-Year')) + 
    scale_linetype_manual(values = c("500-Year" = "solid", "100-Year" = "solid", "25-Year" = "solid"),
                          breaks = c('500-Year', '100-Year', '25-Year'), name = "Return Period") +
  scale_fill_manual(name = "Return Period", values =  c('500-Year' = "#256e3a", '100-Year' = "#5d916c", '25-Year' = "#44c96a"), breaks = c('500-Year', '100-Year', '25-Year'))+
    scale_shape_manual(name = "Return Period", values = c('500-Year' = 15, '100-Year' = 16, '25-Year' = 18), breaks = c('500-Year', '100-Year', '25-Year')) +
     labs(x="", y="", title="Region 3") + theme_minimal() + 
    theme(legend.position="none",  
        axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 15), 
        title = element_text(size = 15))





PARE_plot = reg1_plot + reg2_plot + reg3_plot +plot_layout(ncol = 3) + 
  plot_annotation("Estimated Return Levels - PARE") & theme(plot.title = element_text(size = 15))

#ggsave("Images/PARE_plot.png", plot = PARE_plot, width = 7.5, height = 3.5, units = "in")

PARE_plot

```

### Table of return levels

```{r}
#| label: rl-table

## Table S3
car_rl_25_22_se  <- rl_with_ci(par_mat = par_22, 
                            varcov_list = car_varcov_22, 
                            return_period = 25, 
                            type = "se", alpha = 0.05)/254
car_rl_100_22_se <- rl_with_ci(par_22, car_varcov_22, 100, "se")/254
car_rl_500_22_se <- rl_with_ci(par_22, car_varcov_22, 500, "se")/254

print("Return Levels for First 40 Years of Data (1921-1960)")
rbind(t(car_rl_25_22_se), t(car_rl_100_22_se), t(car_rl_500_22_se))


## latex output
# car_rl_22_list <- list(car_rl_25_22_se, car_rl_100_22_se, car_rl_500_22_se)
# get_latex_table_RL(car_rl_22_list)


## Table S4
car_rl_25_52_se  <- rl_with_ci(par_52, car_varcov_52, 25, "se")/254
car_rl_100_52_se <- rl_with_ci(par_52, car_varcov_52, 100, "se")/254
car_rl_500_52_se <- rl_with_ci(par_52, car_varcov_52, 500, "se")/254

print("Return Levels for Second 40 Years of Data (1951-1990)")
rbind(t(car_rl_25_52_se), t(car_rl_100_52_se), t(car_rl_500_52_se))



# Latex output
# car_rl_52_list <- list(car_rl_25_52_se, car_rl_100_52_se, car_rl_500_52_se)
# get_latex_table_RL(car_rl_52_list)


## TABLE S2 (also Table 3)
car_rl_25_82_se  <- rl_with_ci(par_82, car_varcov_82, 25, "se")/254
car_rl_100_82_se <- rl_with_ci(par_82, car_varcov_82, 100, "se")/254
car_rl_500_82_se <- rl_with_ci(par_82, car_varcov_82, 500, "se")/254

# Latex output
# car_rl_82_list <- list(car_rl_25_82_se, car_rl_100_82_se, car_rl_500_82_se)
# get_latex_table_RL(car_rl_82_list)
print("Return Levels for First 40 Years of Data (1981-2020)")
rbind(t(car_rl_25_82_se), t(car_rl_100_82_se), t(car_rl_500_82_se))
```

## Block Kriging

Data for Block Kriging Return Level plots from [Fagnant '21](https://repository.rice.edu/items/e55295a7-6c2f-499c-be30-6d0132dc9a6e) (Block Kriging sections of Tables 5.3, 5.6, 5.7). Data also included in manuscript in Block Kriging section of Tables S5, S6, S2, and 3.

```{r}
#| label: block_kriging_RL

## read in values for Block Kriging-- read from Fagnant 2021
reg1_bk <- data.frame(RL_25 = c(6.681,7.337,12.008), 
                   RL_100 = c(8.472,9.900,17.277), 
                   RL_500 = c(10.789,13.621,25.800))
reg1_bk$LB_25 = reg1_bk$RL_25 - 1.96*c(1.578,0.670,2.266)
reg1_bk$UB_25 = reg1_bk$RL_25 + 1.96*c(1.578,0.670,2.266)

reg1_bk$LB_100 = reg1_bk$RL_100 - 1.96*c(2.630,1.244,4.485)
reg1_bk$UB_100 = reg1_bk$RL_100 + 1.96*c(2.630,1.244,4.485)

reg1_bk$LB_500 = reg1_bk$RL_500 - 1.96*c(4.319, 2.299,8.969)
reg1_bk$UB_500 = reg1_bk$RL_500 + 1.96*c(4.319, 2.299,8.969)

reg1_bk$LB_X <- c(1957, 1987, 2017)
reg1_bk$UB_X <- c(1963, 1993, 2023)

reg2_bk <- data.frame(RL_25 = c(7.523,6.289,11.210), 
                   RL_100 = c(10.073,7.818,15.465), 
                   RL_500 = c(13.717,9.728,21.911))

reg2_bk$LB_25 = reg2_bk$RL_25 - 1.96*c(1.895, 0.462, 2.06)
reg2_bk$UB_25 = reg2_bk$RL_25 + 1.96*c(1.895, 0.462, 2.06)

reg2_bk$LB_100 = reg2_bk$RL_100 - 1.96*c(3.372,0.771,3.89)
reg2_bk$UB_100 = reg2_bk$RL_100 + 1.96*c(3.372,0.771,3.89)

reg2_bk$LB_500 = reg2_bk$RL_500 - 1.96*c(5.990, 1.260,7.36)
reg2_bk$UB_500 = reg2_bk$RL_500 + 1.96*c(5.990, 1.260,7.36)

reg2_bk$LB_X <- c(1957, 1987, 2017)
reg2_bk$UB_X <- c(1963, 1993, 2023)


reg3_bk <- data.frame(RL_25 = c(7.376,7.950,9.141), 
                   RL_100 = c(9.784,10.650,11.865), 
                   RL_500 = c(13.170,14.511,15.607))

reg3_bk$LB_25 = reg3_bk$RL_25 - 1.96*c(1.885,0.823, 1.44)
reg3_bk$UB_25 = reg3_bk$RL_25 + 1.96*c(1.885,0.823, 1.44)

reg3_bk$LB_100 = reg3_bk$RL_100 - 1.96*c(3.335,1.511, 2.54)
reg3_bk$UB_100 = reg3_bk$RL_100 + 1.96*c(3.335,1.511, 2.54)

reg3_bk$LB_500 = reg3_bk$RL_500 - 1.96*c(5.870,2.757,4.44)
reg3_bk$UB_500 = reg3_bk$RL_500 + 1.96*c(5.870,2.757,4.44)

reg3_bk$LB_X <- c(1957, 1987, 2017)
reg3_bk$UB_X <- c(1963, 1993, 2023)

```

### Visualizations

```{r}
#| label: block-kriging-visuals
## legend 
g <- guides(fill = guide_legend(override.aes = list(color = c("black", "black", "black"),
                                                    shape = c(15, 16, 18), 
                                                    fill = c("black", "black", "black"),
                                                    linetype = c(1,1,1),
                                                    size = c(2,2,3)
                                                    )))

## Region 1
reg1_bk_plot <- ggplot(data=reg1_bk, aes(x=c(1960, 1990, 2020))) + 
    coord_cartesian(ylim = c(3, 41.5)) +
  ## 500 year
    geom_point(aes(y=RL_500, color="500-Year", shape = "500-Year"), size = 2) + 
    geom_errorbar(aes(ymin=LB_500, ymax=UB_500, col="500-Year", linetype = "500-Year"), width = 6)+ 
    geom_rect(aes(ymin=LB_500, ymax=UB_500,xmin = LB_X, xmax = UB_X, fill="500-Year"), alpha=0.1)+
  ## 100 year
    geom_point(aes(y=RL_100, color="100-Year", shape = "100-Year"), size = 2)+ 
    #geom_errorbar(aes(ymin=LB_100, ymax=UB_100, col="100-Year", linetype = "100-Year"), width = 6) +     
    geom_rect(aes(ymin=LB_100, ymax=UB_100,xmin = LB_X, xmax = UB_X, fill="100-Year"), alpha=0.1)+
  ##25 year
    geom_point(aes(y=RL_25, color="25-Year", shape = "25-Year"), size = 3) + 
    geom_errorbar(aes(ymin=LB_25, ymax=UB_25, col="25-Year", linetype = "25-Year"), width = 6) +
    geom_rect(aes(ymin=LB_25, ymax=UB_25,xmin = LB_X, xmax = UB_X, fill="25-Year"), alpha=0.1)+
    scale_x_continuous(breaks = c(1960, 1990,2020)) +
    scale_colour_manual(name = "Return Period", 
                        values = c('500-Year' = "darkblue", '100-Year' = "blue", '25-Year' = "skyblue"),
                        breaks = c('500-Year', '100-Year', '25-Year')) + 
    scale_linetype_manual(values = c("500-Year" = "solid", "100-Year" = "solid", "25-Year" = "solid"),
                          breaks = c('500-Year', '100-Year', '25-Year'), name = "Return Period") +
  scale_fill_manual(name = "Return Period", values =  c('500-Year' = "darkblue", '100-Year' = "blue", '25-Year' = "lightblue"), breaks = c('500-Year', '100-Year', '25-Year'))+
    scale_shape_manual(name = "Return Period", values = c('500-Year' = 15, '100-Year' = 16, '25-Year' = 18), breaks = c('500-Year', '100-Year', '25-Year')) +
     labs(x="", y="Return Level (in)", title="Region 1") + theme_minimal() +   theme(legend.position="none",  
        axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 15), 
        title = element_text(size = 15))


## Region 2
reg2_bk_plot <- ggplot(data=reg2_bk, aes(x=c(1960, 1990, 2020))) + 
    coord_cartesian(ylim = c(3, 41.5)) +
  ## 500 year
    geom_point(aes(y=RL_500, color="500-Year", shape = "500-Year"), size = 2) + 
    geom_errorbar(aes(ymin=LB_500, ymax=UB_500, col="500-Year", linetype = "500-Year"), width = 6)+ 
    geom_rect(aes(ymin=LB_500, ymax=UB_500,xmin = LB_X, xmax = UB_X, fill="500-Year"), alpha=0.1)+
  ## 100 year
    geom_point(aes(y=RL_100, color="100-Year", shape = "100-Year"), size = 2)+ 
    geom_errorbar(aes(ymin=LB_100, ymax=UB_100, col="100-Year", linetype = "100-Year"), width = 6) +     
    geom_rect(aes(ymin=LB_100, ymax=UB_100,xmin = LB_X, xmax = UB_X, fill="100-Year"), alpha=0.1)+
  ##25 year
    geom_point(aes(y=RL_25, color="25-Year", shape = "25-Year"), size = 3) + 
    geom_errorbar(aes(ymin=LB_25, ymax=UB_25, col="25-Year", linetype = "25-Year"), width = 6) +
    geom_rect(aes(ymin=LB_25, ymax=UB_25,xmin = LB_X, xmax = UB_X, fill="25-Year"), alpha=0.1)+
    scale_x_continuous(breaks = c(1960, 1990,2020)) +
    scale_colour_manual(name = "Return Period", 
                        values = c('500-Year' = "darkred", '100-Year' = "red", '25-Year' = "pink"),
                        breaks = c('500-Year', '100-Year', '25-Year')) + 
    scale_linetype_manual(values = c("500-Year" = "solid", "100-Year" = "solid", "25-Year" = "solid"),
                          breaks = c('500-Year', '100-Year', '25-Year'), name = "Return Period") +
  scale_fill_manual(name = "Return Period", values =  c('500-Year' = "darkred", '100-Year' = "red", '25-Year' = "pink"), breaks = c('500-Year', '100-Year', '25-Year'))+
    scale_shape_manual(name = "Return Period", values = c('500-Year' = 15, '100-Year' = 16, '25-Year' = 18), breaks = c('500-Year', '100-Year', '25-Year')) +
     labs(x="Last Year of 40-Year Window", y="", title="Region 2") + theme_minimal() + g +   theme(legend.position="bottom",  
        axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 15), 
        title = element_text(size = 15))



## Region 3
reg3_bk_plot <- ggplot(data=reg3_bk, aes(x=c(1960, 1990, 2020))) + 
    coord_cartesian(ylim = c(3, 41.5)) +
  ## 500 year
    geom_point(aes(y=RL_500, color="500-Year", shape = "500-Year"), size = 2) + 
    geom_errorbar(aes(ymin=LB_500, ymax=UB_500, col="500-Year", linetype = "500-Year"), width = 6)+ 
    geom_rect(aes(ymin=LB_500, ymax=UB_500,xmin = LB_X, xmax = UB_X, fill="500-Year"), alpha=0.1)+
  ## 100 year
    geom_point(aes(y=RL_100, color="100-Year", shape = "100-Year"), size = 2)+ 
    geom_errorbar(aes(ymin=LB_100, ymax=UB_100, col="100-Year", linetype = "100-Year"), width = 6) +     
    geom_rect(aes(ymin=LB_100, ymax=UB_100,xmin = LB_X, xmax = UB_X, fill="100-Year"), alpha=0.1)+
  ##25 year
    geom_point(aes(y=RL_25, color="25-Year", shape = "25-Year"), size = 3) + 
    geom_errorbar(aes(ymin=LB_25, ymax=UB_25, col="25-Year", linetype = "25-Year"), width = 6) +
    geom_rect(aes(ymin=LB_25, ymax=UB_25,xmin = LB_X, xmax = UB_X, fill="25-Year"), alpha=0.1)+
    scale_x_continuous(breaks = c(1960, 1990,2020)) +
    scale_colour_manual(name = "Return Period", 
                        values = c('500-Year' = "#256e3a", '100-Year' = "#5d916c", '25-Year' = "#44c96a"),
                        breaks = c('500-Year', '100-Year', '25-Year')) + 
    scale_linetype_manual(values = c("500-Year" = "solid", "100-Year" = "solid", "25-Year" = "solid"),
                          breaks = c('500-Year', '100-Year', '25-Year'), name = "Return Period") +
  scale_fill_manual(name = "Return Period", values =  c('500-Year' = "#256e3a", '100-Year' = "#5d916c", '25-Year' = "#44c96a"), breaks = c('500-Year', '100-Year', '25-Year'))+
    scale_shape_manual(name = "Return Period", values = c('500-Year' = 15, '100-Year' = 16, '25-Year' = 18), breaks = c('500-Year', '100-Year', '25-Year')) +
     labs(x="", y="", title="Region 3") + theme_minimal() + 
  theme(legend.position="none",  
        axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 15), 
        title = element_text(size = 15))





BK_plot = reg1_bk_plot + reg2_bk_plot + reg3_bk_plot +plot_layout(ncol = 3) + plot_annotation("Estimated Return Levels - Block Kriging") & theme(plot.title = element_text(size = 15))

BK_plot

ggsave("Images/BK_plot.png", plot = BK_plot, width = 7.5, height = 3.5, units = "in")
```

## Regional Max

Data for Regional Max Return Level plots from [Fagnant '21](https://repository.rice.edu/items/e55295a7-6c2f-499c-be30-6d0132dc9a6e) (Regional Max sections of Tables 5.3, 5.6, 5.7). Data also included in manuscript in Regional Max section of Tables S5, S6, S2, and 3.

```{r}
#| label: regional-max-data
## read in values for Regional Max-- read from Fagnant 2021
reg1_rm <- data.frame(RL_25 = c(9.349, 9.552, 12.96), 
                   RL_100 = c(12.073, 12.295, 17.54), 
                   RL_500 = c(15.759, 15.999, 24.23))
reg1_rm$LB_25 = reg1_rm$RL_25 - 1.96*c(1.354, 0.854, 1.33)
reg1_rm$UB_25 = reg1_rm$RL_25 + 1.96*c(1.354, 0.854, 1.33)

reg1_rm$LB_100 = reg1_rm$RL_100 - 1.96*c(2.341, 1.455, 2.39)
reg1_rm$UB_100 = reg1_rm$RL_100 + 1.96*c(2.341, 1.455, 2.39)

reg1_rm$LB_500 = reg1_rm$RL_500 - 1.96*c(4.016, 2.470, 4.32)
reg1_rm$UB_500 = reg1_rm$RL_500 + 1.96*c(4.016, 2.470, 4.32)

reg1_rm$LB_X <- c(1957, 1987, 2017)
reg1_rm$UB_X <- c(1963, 1993, 2023)

reg2_rm <- data.frame(RL_25 = c(10.895, 11.858, 14.68), 
                   RL_100 = c(15.050, 15.272, 19.68), 
                   RL_500 = c(21.297, 19.852, 26.83))

reg2_rm$LB_25 = reg2_rm$RL_25 - 1.96*c(1.283, 1.025, 1.44)
reg2_rm$UB_25 = reg2_rm$RL_25 + 1.96*c(1.283, 1.025, 1.44)

reg2_rm$LB_100 = reg2_rm$RL_100 - 1.96*c(2.402, 1.730, 2.53)
reg2_rm$UB_100 = reg2_rm$RL_100 + 1.96*c(2.402, 1.730, 2.53)

reg2_rm$LB_500 = reg2_rm$RL_500 - 1.96*c(4.517,2.909,4.48)
reg2_rm$UB_500 = reg2_rm$RL_500 + 1.96*c(4.517,2.909,4.48)

reg2_rm$LB_X <- c(1957, 1987, 2017)
reg2_rm$UB_X <- c(1963, 1993, 2023)


reg3_rm <- data.frame(RL_25 = c(10.037,11.636, 13.75), 
                   RL_100 = c(13.553, 16.109, 18.37), 
                   RL_500 = c(18.659, 22.889, 24.94))

reg3_rm$LB_25 = reg3_rm$RL_25 - 1.96*c(1.179, 1.381, 1.35)
reg3_rm$UB_25 = reg3_rm$RL_25 + 1.96*c(1.179, 1.381, 1.35)

reg3_rm$LB_100 = reg3_rm$RL_100 - 1.96*c(2.161,2.578, 2.39)
reg3_rm$UB_100 = reg3_rm$RL_100 + 1.96*c(2.161,2.578, 2.39)

reg3_rm$LB_500 = reg3_rm$RL_500 - 1.96*c(3.956, 4.847, 4.22)
reg3_rm$UB_500 = reg3_rm$RL_500 + 1.96*c(3.956, 4.847, 4.22)

reg3_rm$LB_X <- c(1957, 1987, 2017)
reg3_rm$UB_X <- c(1963, 1993, 2023)

```

### Visualizations

```{r}
#| label: regional-max-visuals
## legend 
g <- guides(fill = guide_legend(override.aes = list(color = c("black", "black", "black"),
                                                    shape = c(15, 16, 18), 
                                                    fill = c("black", "black", "black"),
                                                    linetype = c(1,1,1),
                                                    size = c(2,2,3)
                                                    )))

## Region 1
reg1_rm_plot <- ggplot(data=reg1_rm, aes(x=c(1960, 1990, 2020))) + 
    coord_cartesian(ylim = c(3, 41.5)) +
  ## 500 year
    geom_point(aes(y=RL_500, color="500-Year", shape = "500-Year"), size = 2) + 
    geom_errorbar(aes(ymin=LB_500, ymax=UB_500, col="500-Year", linetype = "500-Year"), width = 6)+ 
    geom_rect(aes(ymin=LB_500, ymax=UB_500,xmin = LB_X, xmax = UB_X, fill="500-Year"), alpha=0.1)+
  ## 100 year
    geom_point(aes(y=RL_100, color="100-Year", shape = "100-Year"), size = 2)+ 
    #geom_errorbar(aes(ymin=LB_100, ymax=UB_100, col="100-Year", linetype = "100-Year"), width = 6) +     
    geom_rect(aes(ymin=LB_100, ymax=UB_100,xmin = LB_X, xmax = UB_X, fill="100-Year"), alpha=0.1)+
  ##25 year
    geom_point(aes(y=RL_25, color="25-Year", shape = "25-Year"), size = 3) + 
    geom_errorbar(aes(ymin=LB_25, ymax=UB_25, col="25-Year", linetype = "25-Year"), width = 6) +
    geom_rect(aes(ymin=LB_25, ymax=UB_25,xmin = LB_X, xmax = UB_X, fill="25-Year"), alpha=0.1)+
    scale_x_continuous(breaks = c(1960, 1990,2020)) +
    scale_colour_manual(name = "Return Period", 
                        values = c('500-Year' = "darkblue", '100-Year' = "blue", '25-Year' = "skyblue"),
                        breaks = c('500-Year', '100-Year', '25-Year')) + 
    scale_linetype_manual(values = c("500-Year" = "solid", "100-Year" = "solid", "25-Year" = "solid"),
                          breaks = c('500-Year', '100-Year', '25-Year'), name = "Return Period") +
  scale_fill_manual(name = "Return Period", values =  c('500-Year' = "darkblue", '100-Year' = "blue", '25-Year' = "lightblue"), breaks = c('500-Year', '100-Year', '25-Year'))+
    scale_shape_manual(name = "Return Period", values = c('500-Year' = 15, '100-Year' = 16, '25-Year' = 18), breaks = c('500-Year', '100-Year', '25-Year')) +
     labs(x="", y="Return Level (in)", title="Region 1") + theme_minimal() +
    theme(legend.position="none",  
        axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 15), 
        title = element_text(size = 15))

## Region 2
reg2_rm_plot <- ggplot(data=reg2_rm, aes(x=c(1960, 1990, 2020))) + 
    coord_cartesian(ylim = c(3, 41.5)) +
  ## 500 year
    geom_point(aes(y=RL_500, color="500-Year", shape = "500-Year"), size = 2) + 
    geom_errorbar(aes(ymin=LB_500, ymax=UB_500, col="500-Year", linetype = "500-Year"), width = 6)+ 
    geom_rect(aes(ymin=LB_500, ymax=UB_500,xmin = LB_X, xmax = UB_X, fill="500-Year"), alpha=0.1)+
  ## 100 year
    geom_point(aes(y=RL_100, color="100-Year", shape = "100-Year"), size = 2)+ 
    geom_errorbar(aes(ymin=LB_100, ymax=UB_100, col="100-Year", linetype = "100-Year"), width = 6) +     
    geom_rect(aes(ymin=LB_100, ymax=UB_100,xmin = LB_X, xmax = UB_X, fill="100-Year"), alpha=0.1)+
  ##25 year
    geom_point(aes(y=RL_25, color="25-Year", shape = "25-Year"), size = 3) + 
    geom_errorbar(aes(ymin=LB_25, ymax=UB_25, col="25-Year", linetype = "25-Year"), width = 6) +
    geom_rect(aes(ymin=LB_25, ymax=UB_25,xmin = LB_X, xmax = UB_X, fill="25-Year"), alpha=0.1)+
    scale_x_continuous(breaks = c(1960, 1990,2020)) +
    scale_colour_manual(name = "Return Period", 
                        values = c('500-Year' = "darkred", '100-Year' = "red", '25-Year' = "pink"),
                        breaks = c('500-Year', '100-Year', '25-Year')) + 
    scale_linetype_manual(values = c("500-Year" = "solid", "100-Year" = "solid", "25-Year" = "solid"),
                          breaks = c('500-Year', '100-Year', '25-Year'), name = "Return Period") +
  scale_fill_manual(name = "Return Period", values =  c('500-Year' = "darkred", '100-Year' = "red", '25-Year' = "pink"), breaks = c('500-Year', '100-Year', '25-Year'))+
    scale_shape_manual(name = "Return Period", values = c('500-Year' = 15, '100-Year' = 16, '25-Year' = 18), breaks = c('500-Year', '100-Year', '25-Year')) +
     labs(x="Last Year of 40-Year Window", y="", title="Region 2") + theme_minimal() + g +   theme(legend.position="bottom",  
        axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 15), 
        title = element_text(size = 15))


## Region 3
reg3_rm_plot <- ggplot(data=reg3_rm, aes(x=c(1960, 1990, 2020))) + 
    coord_cartesian(ylim = c(3, 41.5)) +
  ## 500 year
    geom_point(aes(y=RL_500, color="500-Year", shape = "500-Year"), size = 2) + 
    geom_errorbar(aes(ymin=LB_500, ymax=UB_500, col="500-Year", linetype = "500-Year"), width = 6)+ 
    geom_rect(aes(ymin=LB_500, ymax=UB_500,xmin = LB_X, xmax = UB_X, fill="500-Year"), alpha=0.1)+
  ## 100 year
    geom_point(aes(y=RL_100, color="100-Year", shape = "100-Year"), size = 2)+ 
    geom_errorbar(aes(ymin=LB_100, ymax=UB_100, col="100-Year", linetype = "100-Year"), width = 6) +     
    geom_rect(aes(ymin=LB_100, ymax=UB_100,xmin = LB_X, xmax = UB_X, fill="100-Year"), alpha=0.1)+
  ##25 year
    geom_point(aes(y=RL_25, color="25-Year", shape = "25-Year"), size = 3) + 
    geom_errorbar(aes(ymin=LB_25, ymax=UB_25, col="25-Year", linetype = "25-Year"), width = 6) +
    geom_rect(aes(ymin=LB_25, ymax=UB_25,xmin = LB_X, xmax = UB_X, fill="25-Year"), alpha=0.1)+
    scale_x_continuous(breaks = c(1960, 1990,2020)) +
    scale_colour_manual(name = "Return Period", 
                        values = c('500-Year' = "#256e3a", '100-Year' = "#5d916c", '25-Year' = "#44c96a"),
                        breaks = c('500-Year', '100-Year', '25-Year')) + 
    scale_linetype_manual(values = c("500-Year" = "solid", "100-Year" = "solid", "25-Year" = "solid"),
                          breaks = c('500-Year', '100-Year', '25-Year'), name = "Return Period") +
  scale_fill_manual(name = "Return Period", values =  c('500-Year' = "#256e3a", '100-Year' = "#5d916c", '25-Year' = "#44c96a"), breaks = c('500-Year', '100-Year', '25-Year'))+
    scale_shape_manual(name = "Return Period", values = c('500-Year' = 15, '100-Year' = 16, '25-Year' = 18), breaks = c('500-Year', '100-Year', '25-Year')) +
     labs(x="", y="", title="Region 3") + theme_minimal() + 
  theme(legend.position="none",  
        axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 15), 
        title = element_text(size = 15))


Regmax_plot = reg1_rm_plot + reg2_rm_plot + reg3_rm_plot +plot_layout(ncol = 3) + plot_annotation("Estimated Return Levels - Regional Max") & theme(plot.title = element_text(size = 15))

Regmax_plot

ggsave("Images/Regmax_plot.png", plot = Regmax_plot, width = 7.5, height = 3.5, units = "in")

```
